---
title: "Pre course survey summary"
author: Taavi Päll
output: 
  github_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To better understand the level of R skills and if they have their own dataset for coursework, students were asked to fill in a short online questionnaire before the course start.

Students were asked for answers to these questions:    
- How big is your previous R experience?    
- What operating system and version your computer is running?    
- Did you have running installation of the following software on your computer?    
- Do you have GitHub account?    
- Do you have your own dataset (at least in mind) that you would like to use for individual project?    
- If you answered 'Yes' to previous question, please describe your dataset (name, how many rows/columns/number of variables, are your values categorical or continuous, csv, json, xls, html, etc.)     

Load required functions (libraries):
```{r}
# tidyverse is for data wrangling and plotting
library(tidyverse)
# g-sheets is for data downloading from google form sheet
library(googlesheets)
```

## Import data

Download query spreadsheet from google drive:
```{r}
(my_sheets <- gs_ls())
```

Haha, the first one seems to be what I'm looking for: "Pre-course … "

Identify query sheet:
```{r}
(surv <- gs_title("Pre-course survey (Responses)"))
```

Import data to table (data_frame):
```{r}
(responses <- surv %>% gs_read())
```

Convert wide table to long format (gather):
```{r}
(resp_gathered <- responses %>% 
  gather(key = key, value = value, -Timestamp))
```

## Results

**How big is your previous R experience?**
```{r}
resp_gathered %>% 
  filter(key == "How big is your previous R experience?") %>% 
  ggplot() +
  stat_count(mapping = aes(x = value)) +
  scale_x_discrete(labels = c("Null", "Little", "Have done\nsomething\nuseful")) +
  scale_y_continuous(breaks = seq(from = 0, to = 10, by = 2)) +
  labs(x = "How big is your previous R experience?")
```

Mostly, students declare zero previous experience with R language. Obviously, it seems that five categories that I proposed in questionnaire was too much - three categories would have been sufficient. 

**What operating system and version your computer is running?**

```{r}
os <- resp_gathered %>% 
  filter(key == "What operating system and version your computer is running?") %>% 
  select(value, Timestamp) %>% 
  mutate(value = str_to_lower(value),
         value = str_remove(value, ",.*| x64"),
         os = case_when(
           str_detect(value, "macos") ~ "macOS",
           str_detect(value, "os x") ~ "OS X",
           TRUE ~ "Win"
         ),
         vers = str_extract(value, "[[:digit:][:punct:]]*( pro)?$"),
         os = str_c(os, vers, sep = " "))
os %>% 
  ggplot() +
  stat_count(mapping = aes(x = os))
```

**Did you have running installation of the following software on your computer?**

```{r}
resp_gathered %>% 
  filter(str_detect(key, "Did you have running")) %>% 
  select(value, Timestamp) %>% 
  ggplot() +
  stat_count(mapping = aes(x = value))
```

It's interesting that **git** was installed on all computers. Perhaps the question was ill-posed by me ...

**Do you have GitHub account?**

```{r}
resp_gathered %>% 
  filter(key == "Do you have GitHub account?") %>% 
  select(value, Timestamp) %>% 
  ggplot() +
  stat_count(mapping = aes(x = value)) +
  scale_y_continuous(breaks = seq(from = 0, to = 12, by = 2))
```

Again, mostly all have GitHub account ... but when was that account created?

Let's download user data from GitHub. gh package contains functions to communicate with GitHub API:
```{r}
library(gh)
```

Get members of rstats-tln (this course) organization and extract usernames (login):
```{r}
members <- gh("GET /orgs/rstats-tln/members", .token = Sys.getenv("GITHUB_PAT"))
(username <- map_chr(members, "login"))
```

Download user data:
```{r}
library(glue)
users <- map(username, ~ gh(glue("GET /users/{.x}")))
```

We are interested in creation date of GH user account: 
```{r}
(created_at <- map_chr(users, "created_at"))
```

Plot out user account creation dates on time axis:
```{r}
library(lubridate)
data_frame(username, created_at) %>%
  mutate(created_at = ymd_hms(created_at)) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = created_at), bins = 60) +
  scale_x_datetime()
```

GitHub reveals that **most accounts were created last week...**
